---
title: "ML Data Management Code"
output: html_notebook
---
# ML Data Management Code

## Intro

This is the R Notebook for our ML Data Management Code. We originally worked on this in an RScript, but will now use an R Notebook for it. The data used in this file will come from a simulated dataset.

## Download packages

``` {r include = FALSE}
library(tidyverse) # for data management
library(psych) # to analyze data and descriptive statistics
library(ggplot2) # to make some nice graphs
library(pander) # to make some clean tables
library(gtsummary) # to make some cool summary tables that fit APA format

```

## Load in RScript where functions are created

``` {r include = TRUE}

#source(__.R)

```

## Download Simulated Data

``` {r include = TRUE}

source(simulate-data.R)

```


## Data Management

### Merging data across time points

#### Zach Method A 

Built in studying the use case of needing to bind and merge different datasets that had different elements of missing data in each. This also works to bind datasets together.

``` {r include = TRUE}

# create dummy data
x_pre <- data.frame(p = c(1, 2, 3), a = c(1, 2, 3), b = c(4, 5, 6), d = c(3, 4, 5))
x_post <- data.frame(p = c(1, 2, 3), a = c(5, 2, 9), c = c(4, 5, 6))
y_pre <- data.frame(p = c(4, 5, 6), b = c(1, 2, 3), c = c(4, 5, 6))
y_post <- data.frame(p = c(4, 5, 6), b = c(1, 2, 3), d = c(4, 5, 6), e = c(1, 2, 3))


# make function
my_fun <- function(df){
  df_name <- deparse(substitute(df))
  if (grepl("pre", df_name)) {
    colnames(df)[-1] <- paste(colnames(df)[-1], "t1", sep = "_") 
  }
  else if (grepl("post", df_name)) {
    colnames(df)[-1] <- paste(colnames(df)[-1], "t2", sep = "_")
  }
  return(df)
}

# create a list of data frames
my_list <- list(x_pre = x_pre, x_post = x_post,
                y_pre = y_pre, y_post = y_post)

# for loop through list
for (i in seq_along(my_list)) {
  df_name <- names(my_list)[i]
  if (grepl("pre", df_name)) {
    colnames(my_list[[i]])[-1] <- paste(colnames(my_list[[i]])[-1], "t1", sep = "_") 
  }
  else if (grepl("post", df_name)) {
    colnames(my_list[[i]])[-1] <- paste(colnames(my_list[[i]])[-1], "t2", sep = "_")
  }
}

```

#### Zach Method B

Alternative approach (probably better). Uses anonymous function and map

``` {r include = TRUE}

# re-create the initial list of data frames
my_list <- list(x_pre = x_pre, x_post = x_post,
                y_pre = y_pre, y_post = y_post)

my_list <- Map(
  function(list_df, list_names) {
    if (endsWith(list_names, "pre")) {
      names(list_df)[-1] <- paste(names(list_df)[-1], "t1", sep = "_")
    } 
    else if (endsWith(list_names, "post")) {
      names(list_df)[-1] <- paste(names(list_df)[-1], "t2", sep = "_")
    }
    list_df
  }, 
  my_list, 
  names(my_list)
)

#merge all data frames together
final_product <- my_list %>% 
  reduce(full_join)
  
```


#### Matt Method (before Zach method)

``` {r include = TRUE}

## not shown here, but typically need to clean dataset before merging

merged_data <- data_pre %>% full_join(data_post, by = c("x"), suffix = c("_pre", "_post"))

## not shown here, but typically need to review the merged column and do additional cleaning

```


### Subsetting a dataset

``` {r include = TRUE}

# Selecting certain rows

subset <- data[c(22:60),]

# Selecting certain columns

subset <- data[,c(8:27)]

## This technique can be used to delete a set of columns

data_1 <- dplyr::select(data, -column1:-column18)

# Filtering for a certain response or group

newdata <- data %>% filter(var1 == 1 | var2 == 1)
newdata <- data %>% filter(var1 == 1 & var2 == 1)
newdata <- data %>% filter(var1 == "value" & var2 == "value")

```


### Binding data together from various cohorts / surveys

Matt's very laborious way of binding data together

``` {r include = TRUE}

## Select all of the columns in the datasets to make sure they are all the same 
## and have the same names using dplyr::select(). Then, turn them all into data tables.

data1 <- as.data.table(data1)

## Next, fix all of the variable types to make sure the variable types in each dataset
## is aligned. For example...

data1$v1 <- as.character(data1$v1)
data1$v2 <- as.integer(data1$v2)
data2$v1 <- as.character(data2$v1)
data2$v2 <- as.integer(data2$v2)

## Then, bind the datasets

data <- funion(data1, data2)

```


### Creating scale variables

``` {r include = TRUE}

data <- data %>% 
  mutate(lv1 = ((v1 + v2 + v3) / 3)) %>%
  mutate(lv2 = ((v4 + v5 + v6) / 3)) %>%
  mutate(lv3 = ((v7 + v8 + v9 + v10 + v11) / 5))

```


### Analyzing descriptive statistics of all items to see if there are any issue with direction or values or other factors → data checking guidelines

#### Quick way to look at means and SDs

``` {r include = TRUE}

table <- data[c("v1", "v2", "v3", "v4", "v5")]

table %>% tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})"), 
                       type = list(v1 ~ 'continuous',
                                   v2 ~ 'continuous', 
                                   v3 ~ 'continuous',
                                   v4 ~ 'continuous',
                                   v5 ~ 'continuous'),
                       digits = all_continuous() ~ 2,
                       missing_text = "(Missing)")

```


#### More detailed way to look at descriptives

``` {r include = TRUE}

b1<- summarySE(data, measurevar="scale", na.rm=TRUE)
b2<- summarySE(data, measurevar="v1", na.rm=TRUE)
b3<- summarySE(data, measurevar="v2", na.rm=TRUE)
b4<- summarySE(data, measurevar="v3", na.rm=TRUE)
b5<- summarySE(data, measurevar="v4", na.rm=TRUE)
b6<- summarySE(data, measurevar="v5", na.rm=TRUE)

#Now, we're going to rename the mean (the 3rd variable) to actually be called mean
names(b1)[3] <- "mean"
names(b2)[3] <- "mean"
names(b3)[3] <- "mean"
names(b4)[3] <- "mean"
names(b5)[3] <- "mean"
names(b6)[3] <- "mean"

# Now we are adding a variable that is the item text for what we're looking at
b1$variable <- "Scale"
b2$variable <- "Item 1"
b3$variable <- "Item 2"
b4$variable <- "Item 3"
b5$variable <- "Item 4"
b6$variable <- "Item 5"


# Now we are adding a variable that is the name of what we're looking at
b1$Type <- "Scale"
b2$Type <- "Item 1"
b3$Type <- "Item 2"
b4$Type <- "Item 3"
b5$Type <- "Item 4"
b6$Type <- "Item 5"


# Bind them all together!

b <- rbind(b1,b2,b3,b4,b5,b6)

# Create table

pander(b[c(7,2:6)], caption="Table for Scale and Items",justify="left",round=2)

```


### Handling missing data

``` {r include = TRUE}


```


### Handling duplicates

In this line, we need to identify which variable should not have duplicates to then select it when we remove duplicates

``` {r include = TRUE}

data <- data[!duplicated(data$v1),]

```


### Identifying and removing outliers

``` {r include = TRUE}



```


### Reverse scoring items

``` {r include = TRUE}

data$v1R <- 7 - data$v1
data$v2R <- 7 - data$v2

```

